--- 
title: "Chapter_1" 
author: "Philippe Glaziou, Nim Pathy, Pete Dodd" 
date: "`r Sys.Date()`" 
output:  
  html_fragment: 
    # Don’t include a table of contents 
    toc: no 
    # Set standard figure width to 12 inches 
    fig_width: 12 
    # Don’t write figure captions 
    fig_caption: FALSE 
--- 


```{r setup, include=FALSE} 
# Chapter 1
# Set options, load data, utility functions 

knitr::opts_chunk$set(
  echo = FALSE,
  results = "asis",
  message = FALSE,
  warning = FALSE
)
library(data.table)
library(gtbreport)
library(ggpubr)
library(here)

source(here('report/ch1_2_dataprep.R'))


``` 






`r anch("Fig. 1.1")`

### <span style="color:#F21905">Fig. 1.1</span> Global trend in case notifications of people newly diagnosed with TB, 2016&#8211;2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.1, fig.alt="Global trend in notifications of new and relapse TB cases, 2017-2021"} 

qplot(
  year,
  c.newinc / 1e6,
  data = global[year >= 2017],
  geom = 'line',
  size = I(1)
) +
  xlab('') + ylab('Notifications per year (millions)') +
  theme_gtb()

ggsave(here('report/pdf/fig01.pdf'), width=8.3, height=5.8)
fwrite(global[year>=2017, .(year, c.newinc)], 
       file=here('report/pdf/fig01.csv'))


```









`r anch("Fig. 1.2")`

### <span style="color:#F21905">Fig. 1.2</span> Trends in notifications of new and relapse TB cases by WHO region, 2016&#8211;2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.2, fig.alt="Trends in notifications of new and relapse TB cases by WHO region, 2017-2021"} 

qplot(
  year,
  c.newinc / 1e6,
  data = regional[year >= 2017],
  geom = 'line',
  size = I(1),
  colour = I('dark blue')
) +
  xlab('') + ylab('Notifications per year (millions)') +
  facet_wrap( ~ region, scales = 'free_y') +
  theme_gtb()

ggsave(here('report/pdf/fig02.pdf'), width=12, height=8)
fwrite(regional[year>=2017, .(region, year, c.newinc)], 
       file=here('report/pdf/fig02.csv'))

```









`r anch("Fig. 1.3")`

### <span style="color:#F21905">Fig. 1.3</span> Trends in monthly or quarterly notifications of TB cases in selected high TB burden countries, January 2020&#8211;June 2021, ^a^
<span class="subhead">The black line indicates the average number of monthly or quarterly case notifications in 2019.</span>

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=8, fig_1.3, fig.alt="Trends in monthly or quarterly notifications of TB cases from January 2020 to June 2022, selected high TB burden countries"} 

qplot(variable,
      value,
      data = monthly[(year %in% 2020:2021 & nm == 12)|(year==2022 & !is.na(value))],
      geom = 'line',
      group = year,
      colour = as.character(year),
      size = I(1)) +
  geom_hline(
    aes(yintercept = n2019 / 12),
    linetype = I(1),
    colour = I('#000000'),
    size = I(1)
  ) +
  facet_wrap(~ country, scales = 'free_y') +
  scale_x_continuous(breaks=1:12) +
  xlab('Month, 2020-2022') + ylab('Notifications per month') +
  theme_gtb()

qplot(variable,
      value,
      data = qty[!is.na(value) & year<2022],
      geom = 'line',
      group = year,
      size = I(1),
      colour = as.character(year)) +
  geom_hline(
    aes(yintercept = n2019 / 4),
    linetype = I(1),
    colour = I('#000000'),
    size = I(1)
  ) +
  facet_wrap( ~ country, scales = 'free_y') +
  xlab('Quarter, 2020') + ylab('Notifications per quarter') +
  theme_gtb()

```

<div class="footnote"> ^a^ Data are shown for countries that were able to report provisional national numbers for all months or quarters to WHO by July 2021.</div> 









<!-- `r anch("Fig. 1.5")` -->

<!-- ### <span style="color:#F21905">Fig. 1.5</span> TB notifications in 2020 compared with 2019 in the 30 high TB burden countries -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.5, fig.alt="Shortfaill in TB notifications in the 30 high TB burden countries in 2020 compared with 2019"}  -->

<!-- dta <- est[year==2020 & g.hbc==TRUE, .(iso3, country, shortfall = short * 100)] -->

<!-- ggbarplot( -->
<!--   dta, -->
<!--   x = 'country', -->
<!--   y = 'shortfall', -->
<!--   fill = "#00aaad", -->
<!--   color = 'white', -->
<!--   sort.val = "desc", -->
<!--   xlab = '', -->
<!--   ylab = 'Percentage of TB notifications in 2020 compared with 2019' -->
<!-- ) +  -->
<!--   rotate() +  -->
<!--   geom_hline(yintercept=100) +  -->
<!--   scale_y_continuous(breaks=seq(0,110, by=10)) -->

<!-- ``` -->










<!-- `r anch("Fig. 1.6")` -->

<!-- ### <span style="color:#F21905">Fig. 1.6</span> The 16 countries with the largest contributions to the global shortfall in TB notifications in 2020 compared with 2019 -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.6, fig.alt="The 16 countries with the largest contributions to the global shortfall in TB notifications in 2020 compared with 2019"}  -->

<!-- gshort <- global[year==2019, c.newinc] - global[year==2020, c.newinc] -->
<!-- dta <- -->
<!--   est[year >= yr-1 & -->
<!--         iso3 %in% unique(model$iso3), .(shortfall = -diff(c.newinc)), by=.(iso3, country)] -->
<!-- dta[, pshort := 100 * shortfall / gshort] -->

<!-- # ggplot(data = dta, aes(reorder(country, shortfall), shortfall)) + -->
<!-- #   geom_bar(stat = 'identity', -->
<!-- #            fill = I('#00aaad'), -->
<!-- #            colour = I('black')) + -->
<!-- #   xlab('')  + ylab('Millions (2019)') + -->
<!-- #   coord_flip() + -->
<!-- #   theme_gtb() -->

<!-- ggbarplot( -->
<!--   dta, -->
<!--   x = 'country', -->
<!--   y = 'pshort', -->
<!--   fill = "#00aaad", -->
<!--   color = 'white', -->
<!--   sort.val = "asc", -->
<!--   xlab = '', -->
<!--   ylab = 'Percentage of global shortfall in 2020 TB notifications compared with 2019' -->
<!-- ) +  -->
<!--   rotate() -->

<!-- ``` -->









`r anch("Fig. 1.7")`

### <span style="color:#F21905">Fig. 1.7</span> Estimated impact of the COVID-19 pandemic on TB mortality for 31 selected countries, 1990-2021

<span class="subhead">TB mortality rate (including HIV)^a^</span> 


```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=11, fig_1.7, fig.alt="Estimated impact of the COVID-19 pandemic on TB mortality for 16 selected countries, up to 2025"} 


qplot(
  as.Date(date),
  best,
  data = model2[measure == 'mort' &
                 hiv == 'a' & year<2022],
  colour = scenario,
  size = I(1),
  geom = 'line'
) +
  xlab('') + ylab('TB mortality rate per 100 000 population per month') +
  facet_wrap( ~ country, scales = 'free_y') +
  scale_colour_manual(values = c("black", "red"))+
  theme_gtb() +
  theme(legend.position = 'none')

```

<!-- <div class="footnote"> ^a^ These estimates are standardized so that rates in January 2020 equal 100 and all subsequent rates are relative to January 2019. For example, a reading of 115 translates into a 15% increase relative to January 2019. Baseline is a scenario of no COVID-19 disruptions based on pre-2020 trends. The impact of COVID-19 related disruptions on estimated mortality is noticeable from 2020 onward.</div>  -->







`r anch("Fig. 1.8")`

### <span style="color:#F21905">Fig. 1.8</span> Estimated impact of the COVID-19 pandemic on TB incidence for 16 selected countries, 2019-2021

<span class="subhead">TB incidence rate^a^</span> 


```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=11, fig_1.8, fig.alt="Estimated impact of the COVID-19 pandemic on TB incidence for 16 selected countries, up to 2025"} 

qplot(
  as.Date(date),
  best,
  data = model2[measure == 'inc' &
                 hiv == 'a' & year<2022],
  colour = scenario,
  size = I(1),
  geom = 'line'
) +
  xlab('') + ylab('TB incidence rate per 100 000 population per month') +
  facet_wrap( ~ country) +
  scale_colour_manual(values = c("black", "blue"))+
#  labs(title = bquote('Standardized TB incidence' ~rate^a)) +
  scale_y_log10() +
  theme_gtb() +
  theme(legend.position = 'none')

```

<!-- <div class="footnote">^a^ These estimates are standardized so that rates in January 2019 equal 100 and all subsequent rates are relative to January 2019. For example, a reading of 115 translates into a 15% increase relative to January 2019. Baseline is a counterfactual scenario of no COVID-19, based on pre-2020 trends. The impact of COVID-19 related disruptions on estimated incidence is limited in 2020 and more noticeable in subsequent years.</div>  -->







