--- 
title: "Section 2.2 TB mortality - interactive charts" 
author: "Takuya Yamanaka" 
date: "`r Sys.Date()`" 
knit: (function(inputFile, encoding) {
  out_dir <- "./html_drafts";
  rmarkdown::render(inputFile,
                    encoding=encoding,
                    output_dir=file.path(dirname(inputFile), out_dir))})
output:  
  html_fragment: 
    # Don’t include a table of contents 
    toc: no 
    # Set standard figure width to 12 inches 
    fig_width: 12 
    # Don’t write figure captions 
    fig_caption: FALSE 
    # Don't embed external resources (stylesheets, JS libraries) in the output 
    self_contained: FALSE    

# To run this file and store output as html:
# rmarkdown::render(here::here("report/ch2-2_kendo.rmd"), output_file = "ch2-2_kendo.html", output_dir = here::here("report/html_drafts/"))
--- 



```{r setup, include=FALSE} 
# Chapter 1
# Set options, load data, utility functions 

knitr::opts_chunk$set(echo = FALSE,  
                      results = "asis", 
                      message = FALSE, 
                      warning = FALSE) 
library(data.table)
library(gtbreport)
library(here)
library(kableExtra)
library(gridExtra)
library(whomap)
library(ggpubr)
library(productplots)
library(RColorBrewer)
library(tidyverse)

library(jsonlite)

source(here('report/ch1_2_dataprep.R'))


``` 


```{css, echo=FALSE}

/* Styles to make it easier to see in the html_fragment; this CSS can be included in the CSS widget of Sitefinity */
  

.red, .red a {
  color: #F21905; /* red text to show figure number */
}


```

<!-- Add links to stylesheets, JQuery and Kendo UI  -->

<link rel="stylesheet" href="https://kendo.cdn.telerik.com/2022.1.301/styles/kendo.common.min.css" />
<link rel="stylesheet" href="https://kendo.cdn.telerik.com/2022.1.301/styles/kendo.default.min.css" />
<link rel="stylesheet" href="https://kendo.cdn.telerik.com/2022.1.301/styles/kendo.default.mobile.min.css" />
<link rel="stylesheet" href="https://maxcdn.bootstrapcdn.com/bootstrap/3.4.1/css/bootstrap.min.css">

<script src="https://kendo.cdn.telerik.com/2022.1.301/js/jquery.min.js"></script>
<script src="https://kendo.cdn.telerik.com/2022.1.301/js/kendo.all.min.js"></script>
<script src="https://docs.telerik.com/kendo-ui/assets/bootstrap.min.js"></script>
<!-- End of links section -->


## 2.2 TB mortality

_Draft! Prepared `r Sys.Date()`_


### <span style="color:#F21905">Table 2.2.1</span> Global and regional estimates of TB mortality, numbers (in thousands) and rates (per 100 000 population) in 2021

<span class="red">**!static figure!**</span>



### <span style="color:#F21905">Fig. 2.2.1</span> Global trends in the estimated number of TB deaths and the mortality rate, 2000&#8211;2021


<div class="col-md-6">
#### (a) Number
<div id="fig_2_2_1a"></div>
</div>
<div class="col-md-6">
#### (b) Rate per 100 000 population
<div id="fig_2_2_1b"></div>
</div>




### <span style="color:#F21905">Fig. 2.2.2</span> Global distribution of estimated TB mortality in HIV-negative people by age group and sex (female in <span style="color:#951b81">purple</span>; male in <span style="color:#80a51b">green</span>), 2021 

<span class="red">**!static figure!**</span>


### <span style="color:#F21905">Fig. 2.2.3</span> Top causes of death worldwide in 2019^a,b^


<div id="fig_2_2_3"></div>




### <span style="color:#F21905">Fig. 2.2.4</span> Estimated number of deaths from HIV/AIDS and TB in 2021^a,b^ 


<div id="fig_2_2_4"></div>




### <span style="color:#F21905">Fig. 2.2.5</span> Global trends in the estimated number of deaths caused by TB and HIV (in millions), 2000&#8211;2021^a,b^

<div id="fig_2_2_5"></div>





### <span style="color:#F21905">Fig. 2.2.6</span> Trends in estimated TB mortality rates by WHO region, 2000&#8211;2021



<div class="row">
<div class="col-md-4">
<div id="fig_2_2_6_afro"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_6_amro"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_6_searo"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_6_euro"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_6_emro"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_6_wpro"></div>
</div>
</div>




### <span style="color:#F21905">Fig. 2.2.7</span> Trends in the estimated absolute number of TB deaths (HIV-positive and HIV-negative, in thousands) by WHO region, 2000&#8211;2021 


<div class="row">
<div class="col-md-4">
<div id="fig_2_2_7_afro"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_7_amro"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_7_searo"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_7_euro"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_7_emro"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_7_wpro"></div>
</div>
</div>


### <span style="color:#F21905">Fig. 2.2.8</span> Regional distribution of estimated TB mortality in HIV-negative people by age group and sex (female in <span style="color:#951b81">purple</span>; male in <span style="color:#80a51b">green</span>), 2021

<span class="red">**!static figure!**</span>



### <span style="color:#F21905">Fig. 2.2.9</span> Trends in the estimated absolute number (in thousands) of TB deaths (HIV-positive and HIV-negative TB) in the 30 high TB burden countries, 2000&#8211;2021

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_9_AGO"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_BGD"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_BRA"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_9_CAF"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_CHN"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_COG"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_9_PRK"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_COD"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_ETH"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_9_GAB"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_IND"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_IDN"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_9_KEN"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_LSO"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_LBR"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_9_MNG"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_MOZ"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_MMR"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_9_NAM"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_NGA"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_PAK"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_9_PNG"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_PHL"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_SLE"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_9_ZAF"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_THA"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_UGA"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_9_TZA"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_VNM"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_9_ZMB"></div>
</div>
</div>


### <span style="color:#F21905">Fig. 2.2.10</span> Trends in the estimated absolute number (in thousands) of TB deaths (HIV-positive and HIV-negative TB) in the 3 global TB watchlist countries, 2000&#8211;2021

<div class="row">
<div class="col-md-4">
<div id="fig_2_2_10_KHM"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_10_RUS"></div>
</div>
<div class="col-md-4">
<div id="fig_2_2_10_ZWE"></div>
</div>
</div>


### <span style="color:#F21905">Fig. 2.2.11</span> Countries which, by 2021, had reached the 2020 milestone of the End TB Strategy for reducing the total number of TB deaths

<span class="red">**!static figure!**</span>



### <span style="color:#F21905">Fig. 2.2.12</span> Estimated TB mortality rates in HIV-negative people, 2021

<span class="red">**!static figure!**</span>




### <span style="color:#F21905">Fig. 2.2.13</span> Estimates of the case fatality ratio (CFR), including HIV-negative and HIV-positive people, 2021

<span class="red">**!static figure!**</span>





<script type="text/javascript">

/* JSON data objects for the figures */

```{r, json_2_1} 
# renaming variable for Kendo UI
names(global) <- gsub('[.]', '_', names(global))

names(regional) <- gsub('[.]', '_', names(regional))

names(hest) <- gsub('[.]', '_', names(hest))

# fig 2.2.1
dta_2_2_1 <- est[iso3 %in% c('KHM','RUS','ZWE')]
dta_2_2_1[, inc.milestone := inc[16] * 0.8, by = iso3]
names(dta_2_2_1) <- gsub('[.]', '_', names(dta_2_2_1))

# fig 2.2.3
tbhiv <- global$mort_h_num[global$year == 2019] / 1e3
top10$tbhiv <- c(rep(0, 12), tbhiv, rep(0, 7))
top10$n <- top10$deaths / 1000 + top10$tbhiv / 1000

# fig 2.2.4
names(unaids) <- gsub('[.]', '_', names(unaids))
hivd <- unaids[, .(sum(mort_hiv_num, na.rm = T)), by = year]
cod <- data.table(
  cause = c('HIV/AIDS', 'TB'),
  n = c((last(hivd$V1) - last(global$mort_h_num)) , last(global$mort_nh_num) ),
  tbhiv = rep(last(global$mort_h_num) , 2)
)
cod[, total := n + tbhiv]

# fig 2.2.5
m <- 1e5
ghiv.mort <-
  est[, addXY(mort.hiv / m, r.sd = mort.hiv.sd / m, weights = as.numeric(pop)), by =
        year]
dta_2_2_5 <- merge(global, ghiv.mort, by = 'year')
names(dta_2_2_5) <- gsub('[.]', '_', names(dta_2_2_5))

# fig 2.2.10
dta_2_2_10 <- est[iso3 %in% c('KHM','RUS','ZWE')]
dta_2_2_10[, mort.milestone := mort.num[16] * 0.65, by = iso3]
names(dta_2_2_10) <- gsub('[.]', '_', names(dta_2_2_10))

```

var fig_2_2_1a_data = `r global %>% select(year,mort_num,mort_lo_num,mort_hi_num,mort_h_num,mort_h_hi_num,mort_h_lo_num,mort_nh_num,mort_nh_hi_num,mort_nh_lo_num) %>% mutate(milestone = as.numeric(select(filter(global,year==2015),mort_num))*0.65) %>% toJSON("rows")`; 
var fig_2_2_1b_data = `r global %>% select(year,mort,mort_lo,mort_hi,mort_h,mort_h_hi,mort_h_lo,mort_nh,mort_nh_hi,mort_nh_lo)  %>% toJSON("rows")`; 

var fig_2_2_3_data = `r top10  %>% select(cause,deaths,tbhiv) %>% toJSON("rows")`; 

var fig_2_2_4_data = `r cod %>% arrange(desc(n)) %>% toJSON("rows")`; 

var fig_2_2_5_data = `r dta_2_2_5 %>% select(year,mort_h_num,mort_h_hi_num,mort_h_lo_num,mort_nh_num,mort_nh_hi_num,mort_nh_lo_num,r_num,r_lo_num,r_hi_num) %>% toJSON("rows")`; 

var fig_2_2_6_data = `r regional %>% select(region,year,mort_h,mort_h_hi,mort_h_lo,mort_nh,mort_nh_hi,mort_nh_lo)  %>% toJSON("rows")`; 

var fig_2_2_7_data = `r regional %>% select(entity = region,year,mort_num,mort_hi_num,mort_lo_num,mort_milestone)  %>% toJSON("rows")`; 

var fig_2_2_9_data = `r hest %>% select(entity = country,year,mort_num,mort_hi_num,mort_lo_num,mort_milestone)  %>% toJSON("rows")`; 

var fig_2_2_10_data = `r dta_2_2_10 %>% select(entity = country,year,mort_num,mort_hi_num,mort_lo_num,mort_milestone)  %>% toJSON("rows")`; 

</script>


```{js, echo=FALSE}

/* Functions to create the figures */

/* Functions to organize number with thousands separator */
function num_spacer(number) {
  // Change thousands separator to a non-breaking space as per WHO standard
      if (number >= 10 | number==0) {
        return kendo.toString(number, 'n0').replace(/,/g, '&nbsp;');
      } else if (number<10 & number>=1) {
      return kendo.toString(number, 'n1').replace(/,/g, '&nbsp;');
      } else {
      return kendo.toString(number, 'n2').replace(/,/g, '&nbsp;');
      }
}

function axis_spacer(number) {
  // Change thousands separator to a non-breaking space as per WHO standard
      if (number >= 10 | number == 0 ) {
        return kendo.toString(number, 'n0').replace(/,/g, ' ');
      } else {return kendo.toString(number, 'n1').replace(/,/g, ' ');
             }
}

/* format a number as thousands using GTB rounding rules */
function tb_format_thou(n) { 
  nt = n/1000; 
  if (nt < 100) {
    ntform = Number(nt.toPrecision(2))*1000;
      } else {
    ntform = Number(nt.toPrecision(3))*1000;
  } 
  return num_spacer(ntform)
}


function createFig_2_2_1a(fig_ID) {
 
		$(fig_ID).kendoChart({
			dataSource: fig_2_2_1a_data,
			chartArea: {
				height: 500
			},	
			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
        name: "Total number TB deaths",
				field: "mort_num",
				color: "black",
				tooltip: {
					visible: true,
					template: "Total number of TB deaths (#= category #): #= Number((value/1e6).toPrecision(2)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "mort_lo_num",
				toField: "mort_hi_num",
				color: "black",
				tooltip: {
					visible: true,
  				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(2)) #–#= Number((value.to/1e6).toPrecision(2)) #"
				}
			},{
				type: "line",
        name: "TB deaths in HIV-positive people",
				field: "mort_h_num",
				color: "lightskyblue",
				tooltip: {
					visible: true,
					template: "TB deaths in HIV-positive people (#= category #): #= Number((value/1e6).toPrecision(2)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "mort_h_lo_num",
				toField: "mort_h_hi_num",
				color: "lightskyblue",
				tooltip: {
					visible: true,
  				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(2)) #–#= Number((value.to/1e6).toPrecision(2))  #"
				}
			},{
				type: "line",
        name: "TB deaths in HIV-negative people",
				field: "mort_nh_num",
				color: "blue",
				tooltip: {
					visible: true,
					template: "TB deaths in HIV-negative people (#= category #): #= Number((value/1e6).toPrecision(2)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "mort_nh_lo_num",
				toField: "mort_nh_hi_num",
				color: "blue",
				tooltip: {
					visible: true,
  				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(2)) #–#= Number((value.to/1e6).toPrecision(2)) #"
				}
			},{
        type: "line",
        dashType: "dash",
        field: "milestone",
        color: "grey",
        markers: {
             size: 0,
             opacity: 0
           },
			tooltip: {
				visible: false
			}
      },],      
            render: function (e) {
            var draw = kendo.drawing;
            var padding = 100;
            var padding2 = 145;
            var element = this.element;
            var rect = new kendo.geometry.Rect([padding, padding2], [element.width() - 2 * padding, element.height() - 2 * padding2]);

            var text = new draw.Text("2020 milestone", [0, 0], { font: "bold 9px Verdana,Arial,sans-serif" });
            draw.align([text], rect, "start");
            draw.vAlign([text], rect, "start");
            e.sender.surface.draw(text);
          },
			valueAxis: {
				type: "log",
        labels: {
					template: "#= axis_spacer(value/1e6) #"
				},
				title: {
					text: "Millions per year (log scale)",
					visible: true
				},
        min: 1e5,
        majorUnit: 2,
				line: {
					visible: false
				}
			},
			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 5
				},
				majorGridLines: {
					visible: false
				},
				title: {
					text: "Year",
					visible: true
				}
			}

		});
}



function createFig_2_2_1b(fig_ID) {
 
		$(fig_ID).kendoChart({
			dataSource: fig_2_2_1b_data,
			chartArea: {
				height: 500
			},	
			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
        name: "Total TB deaths",
				field: "mort",
				color: "black",
				tooltip: {
					visible: true,
					template: "Total TB deaths (#= category #): #= tb_format_thou(value) # per 100&nbsp;000 population"
				}
			},{
				type: "rangeArea",
				fromField: "mort_lo",
				toField: "mort_hi",
				color: "black",
				tooltip: {
					visible: true,
  				template: "95% uncertainty interval (#= category #): #= tb_format_thou(value.from) #–#= tb_format_thou(value.to) #"
				}
			},{
				type: "line",
        name: "TB deaths in HIV-positive people",
				field: "mort_h",
				color: "lightskyblue",
				tooltip: {
					visible: true,
					template: "TB deaths in HIV-positive people (#= category #): #= tb_format_thou(value) # per 100&nbsp;000 population"
				}
			},{
				type: "rangeArea",
				fromField: "mort_h_lo",
				toField: "mort_h_hi",
				color: "lightskyblue",
				tooltip: {
					visible: true,
				  template: "95% uncertainty interval (#= category #): #= tb_format_thou(value.from)  #–#= tb_format_thou(value.to)  #"
				}
			},{
				type: "line",
        name: "TB deaths in HIV-negative people",
				field: "mort_nh",
				color: "blue",
				tooltip: {
					visible: true,
					template: "TB deaths in HIV-negative people (#= category #): #= tb_format_thou(value) # per 100&nbsp;000 population"
				}
			},{
				type: "rangeArea",
				fromField: "mort_nh_lo",
				toField: "mort_nh_hi",
				color: "blue",
				tooltip: {
					visible: true,
				  template: "95% uncertainty interval (#= category #): #= tb_format_thou(value.from) #–#= tb_format_thou(value.to) #"
				}
			},],      

			valueAxis: {
				type: "log",
        labels: {
					template: "#= axis_spacer(value) #"
				},
				title: {
					text: "Rate per 100 000 population per year (log scale)",
					visible: true
				},
        min: 1,
        majorUnit: 2,
				line: {
					visible: false
				}
			},
			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 5
				},
				majorGridLines: {
					visible: false
				},
				title: {
					text: "Year",
					visible: true
				}
			}

		});
}

function createFig_2_2_3(fig_ID) {
		$(fig_ID).kendoChart({
			dataSource: fig_2_2_3_data,
			chartArea: {
				height: 600
			},	
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "bar",
        stack: true
			},
			series: [{
				field: "deaths",
				color: "#00aaad",
        gap: 0.2,
        tooltip: {
				visible: true,
				template: "#= category #: #= Number((value/1e3).toPrecision(2)) # million"
			}
			},{
				field: "tbhiv",
				color: "grey",
        tooltip: {
				visible: true,
				template: "#= category #, TB-HIV: #= Number((value/1e3).toPrecision(2)) # million"
			}
			},],
			valueAxis: {
				labels: {
					template: "#= axis_spacer(value/1e3) #"
				},
				title: {
					text: "Number of deaths (millions)"
				},
				line: {
					visible: false
				}
			},
			categoryAxis: {
				field: "cause",
				labels: {
					rotation: "auto"
				},
				majorGridLines: {
					visible: false
				}
			},

		});
}



function createFig_2_2_4(fig_ID) {
		$(fig_ID).kendoChart({
			dataSource: fig_2_2_4_data,
			chartArea: {
				height: 300
			},	
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "bar",
        stack: true
			},
			series: [{
				field: "n",
				color: "#00aaad",
        gap: 0.2,
        tooltip: {
				visible: true,
				template: "#= category #: #= Number((value/1e6).toPrecision(2)) # million"
			}
			},{
				field: "tbhiv",
				color: "grey",
        tooltip: {
				visible: true,
				template: "TB-HIV: #= Number((value/1e6).toPrecision(2)) # million"
			}
			},],
			valueAxis: {
				labels: {
					template: "#= num_spacer(value/1e6) #"
				},
				title: {
					text: "Number of deaths (millions)"
				},
				line: {
					visible: false
				}
			},
			categoryAxis: {
				field: "cause",
				labels: {
					rotation: "auto"
				},
				majorGridLines: {
					visible: false
				}
			},

		});
}

function createFig_2_2_5(fig_ID) {
 
		$(fig_ID).kendoChart({
			dataSource: fig_2_2_5_data,
			chartArea: {
				height: 500
			},	
			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
        name: "HIV deaths",
				field: "r_num",
				color: "black",
				tooltip: {
					visible: true,
					template: "HIV deaths (#= category #): #= Number((value/1e6).toPrecision(2)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "r_lo_num",
				toField: "r_hi_num",
				color: "black",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(2))#–#= Number((value.to/1e6).toPrecision(2)) #"
				}
			},{
				type: "line",
        name: "TB deaths in HIV-positive people",
				field: "mort_h_num",
				color: "lightskyblue",
				tooltip: {
					visible: true,
					template: "TB deaths in HIV-positive people (#= category #): #= Number((value/1e6).toPrecision(2)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "mort_h_lo_num",
				toField: "mort_h_hi_num",
				color: "lightskyblue",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(2))  #–#= Number((value.to/1e6).toPrecision(2))  #"
				}
			},{
				type: "line",
        name: "TB deaths in HIV-negative people",
				field: "mort_nh_num",
				color: "blue",
				tooltip: {
					visible: true,
					template: "TB deaths in HIV-negative people (#= category #): #= Number((value/1e6).toPrecision(2)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "mort_nh_lo_num",
				toField: "mort_nh_hi_num",
				color: "blue",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(2)) #–#= Number((value.to/1e6).toPrecision(2)) #"
				}
			},],      

			valueAxis: {
				type: "log",
        labels: {
					template: "#= axis_spacer(value/1e6) #"
				},
				title: {
					text: "Millions of deaths per year (log scale)",
					visible: true
				},
        min: 1e5,
        majorUnit: 2,
				line: {
					visible: false
				}
			},
			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 5
				},
				majorGridLines: {
					visible: false
				},
				title: {
					text: "Year",
					visible: true
				}
			}

		});
}


function createFig_2_2_6(fig_ID, data, filter, min, unit) {
 
  	// Filter the dataset on the country variable
		dataJSON = data.filter( element => element.region == filter);
  
		$(fig_ID).kendoChart({		
      dataSource: dataJSON,			
			chartArea: {
				height: 400
			},	      
      title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},	  

			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
				field: "mort_nh",
				color: "blue",
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "TB mortality rate among HIV-negative people (#= category #): #= tb_format_thou(value) #"
				}
			},{
				type: "rangeArea",
				fromField: "mort_nh_lo",
				toField: "mort_nh_hi",
				color: "blue",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= tb_format_thou(value.from) #–#= tb_format_thou(value.to) #"
				}
			},{
				type: "line",
				field: "mort_h",
				color: "lightskyblue", 
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "TB mortality rate among HIV-positive people (#= category #): #= tb_format_thou(value) #"
				}
			},{
				type: "rangeArea",
				fromField: "mort_h_lo",
				toField: "mort_h_hi",
				color: "lightskyblue",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= tb_format_thou(value.from) #–#= tb_format_thou(value.to) #"
				}
			},],      
      
			valueAxis: {
				type: "log",
        labels: {
					template: "#= axis_spacer(value) #"
				},
				title: {
					text: "Mortality rate per 100 000 population\nper year (log scale)",
					visible: true
				},
       min: min,
        majorUnit: unit,
        axisCrossingValue: 0.1,
				line: {
					visible: false
				}
			},      

			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 5
				},
				majorGridLines: {
					visible: false
				},

				title: {
					text: "Year",
					visible: true
				}
			}

		});
}
 
function createFig_2_2_7910(fig_ID, data, filter, fig_height, min, unit) {
 
  	// Filter the dataset on the entity variable
		dataJSON = data.filter( element => element.entity == filter);
  
		$(fig_ID).kendoChart({		
      dataSource: dataJSON,			
			chartArea: {
				height: fig_height
			},	      
      title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},	  

			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
				field: "mort_num",
				color: "black",
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "Total TB deaths (#= category #): #= Number((value/1e3).toPrecision(2)) # thousand"
				}
			},{
				type: "rangeArea",
				fromField: "mort_lo_num",
				toField: "mort_hi_num",
				color: "black",
				tooltip: {
					visible: true,
				  template: "95% uncertainty interval (#= category #): #= Number((value.from/1e3).toPrecision(2)) #–#= Number((value.to/1e3).toPrecision(2)) #"
				}
			},{
        type: "line",
        dashType: "dash",
        field: "mort_milestone",
        color: "grey",
        markers: {
             size: 0,
             opacity: 0
           },
			tooltip: {
				visible: false
			}
      },],      
      
			valueAxis: {
				type: "log",
        labels: {
					template: "#= axis_spacer(value/1e3) #"
				},
				title: {
					text: "Total TB deaths per year\n(thousands, log scale)",
					visible: true
				},
       min: min,
        majorUnit: unit,
        axisCrossingValue: min,
				line: {
					visible: false
				}
			},      

			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 5
				},
				majorGridLines: {
					visible: false
				},

				title: {
					text: "Year",
					visible: true
				}
			}

		});
}


```


```{js, echo=FALSE}

$(document).ready(function () {
createFig_2_2_1a("#fig_2_2_1a");  
createFig_2_2_1b("#fig_2_2_1b");
createFig_2_2_3("#fig_2_2_3");
createFig_2_2_4("#fig_2_2_4");
createFig_2_2_5("#fig_2_2_5");                 
createFig_2_2_6("#fig_2_2_6_euro",fig_2_2_6_data,"European Region",0.1,10);
createFig_2_2_6("#fig_2_2_6_afro",fig_2_2_6_data,"African Region",1,10);
createFig_2_2_6("#fig_2_2_6_emro",fig_2_2_6_data,"Eastern Mediterranean Region",0.1,30);
createFig_2_2_6("#fig_2_2_6_amro",fig_2_2_6_data,"Region of the Americas",0.1,4);
createFig_2_2_6("#fig_2_2_6_wpro",fig_2_2_6_data,"Western Pacific Region",0.1,20);
createFig_2_2_6("#fig_2_2_6_searo",fig_2_2_6_data,"South-East Asia Region",0.1,80); 

createFig_2_2_7910("#fig_2_2_7_euro",fig_2_2_7_data,"European Region",400,19e3,1.4);
createFig_2_2_7910("#fig_2_2_7_afro",fig_2_2_7_data,"African Region",400,300e3,1.4);
createFig_2_2_7910("#fig_2_2_7_emro",fig_2_2_7_data,"Eastern Mediterranean Region",400,50e3,1.4);
createFig_2_2_7910("#fig_2_2_7_amro",fig_2_2_7_data,"Region of the Americas",400,15e3,1.4);
createFig_2_2_7910("#fig_2_2_7_wpro",fig_2_2_7_data,"Western Pacific Region",400,60e3,1.4);
createFig_2_2_7910("#fig_2_2_7_searo",fig_2_2_7_data,"South-East Asia Region",400,400e3,1.4);

createFig_2_2_7910("#fig_2_2_9_AGO",fig_2_2_9_data,"Angola",250,5e3,1.8);
createFig_2_2_7910("#fig_2_2_9_COG",fig_2_2_9_data,"Congo",250,1e3,2);
createFig_2_2_7910("#fig_2_2_9_IND",fig_2_2_9_data,"India",250,250e3,2);
createFig_2_2_7910("#fig_2_2_9_MNG",fig_2_2_9_data,"Mongolia",250,0.2e3,1.5);
createFig_2_2_7910("#fig_2_2_9_PAK",fig_2_2_9_data,"Pakistan",250,30e3,1.4);
createFig_2_2_7910("#fig_2_2_9_THA",fig_2_2_9_data,"Thailand",250,7e3,2);

createFig_2_2_7910("#fig_2_2_9_BGD",fig_2_2_9_data,"Bangladesh",250,20e3,1.8);
createFig_2_2_7910("#fig_2_2_9_PRK",fig_2_2_9_data,"Democratic People's Republic of Korea",250,5e3,2);
createFig_2_2_7910("#fig_2_2_9_IDN",fig_2_2_9_data,"Indonesia",250,60e3,1.4);
createFig_2_2_7910("#fig_2_2_9_MOZ",fig_2_2_9_data,"Mozambique",250,5e3,2);
createFig_2_2_7910("#fig_2_2_9_PNG",fig_2_2_9_data,"Papua New Guinea",250,2e3,2);
createFig_2_2_7910("#fig_2_2_9_UGA",fig_2_2_9_data,"Uganda",250,8e3,1.7);

createFig_2_2_7910("#fig_2_2_9_BRA",fig_2_2_9_data,"Brazil",250,4e3,1.6);
createFig_2_2_7910("#fig_2_2_9_COD",fig_2_2_9_data,"Democratic Republic of the Congo",250,25e3,1.6);
createFig_2_2_7910("#fig_2_2_9_KEN",fig_2_2_9_data,"Kenya",250,15e3,2);
createFig_2_2_7910("#fig_2_2_9_MMR",fig_2_2_9_data,"Myanmar",250,10e3,2);
createFig_2_2_7910("#fig_2_2_9_PHL",fig_2_2_9_data,"Philippines",250,15e3,1.6);
createFig_2_2_7910("#fig_2_2_9_TZA",fig_2_2_9_data,"United Republic of Tanzania",250,10e3,2);

createFig_2_2_7910("#fig_2_2_9_CAF",fig_2_2_9_data,"Central African Republic",250,2e3,2);
createFig_2_2_7910("#fig_2_2_9_ETH",fig_2_2_9_data,"Ethiopia",250,10e3,2);
createFig_2_2_7910("#fig_2_2_9_LSO",fig_2_2_9_data,"Lesotho",250,3e3,2);
createFig_2_2_7910("#fig_2_2_9_NAM",fig_2_2_9_data,"Namibia",250,1e3,2);
createFig_2_2_7910("#fig_2_2_9_SLE",fig_2_2_9_data,"Sierra Leone",250,2e3,2);
createFig_2_2_7910("#fig_2_2_9_VNM",fig_2_2_9_data,"Viet Nam",250,5e3,2);

createFig_2_2_7910("#fig_2_2_9_CHN",fig_2_2_9_data,"China",250,25e3,1.6);
createFig_2_2_7910("#fig_2_2_9_GAB",fig_2_2_9_data,"Gabon",250,0.5e3,2);
createFig_2_2_7910("#fig_2_2_9_LBR",fig_2_2_9_data,"Liberia",250,1e3,2);
createFig_2_2_7910("#fig_2_2_9_NGA",fig_2_2_9_data,"Nigeria",250,60e3,1.4);
createFig_2_2_7910("#fig_2_2_9_ZAF",fig_2_2_9_data,"South Africa",250,20e3,2);
createFig_2_2_7910("#fig_2_2_9_ZMB",fig_2_2_9_data,"Zambia",250,4e3,2);

createFig_2_2_7910("#fig_2_2_10_KHM",fig_2_2_10_data,"Cambodia",250,2e3,2);
createFig_2_2_7910("#fig_2_2_10_RUS",fig_2_2_10_data,"Russian Federation",250,5e3,2);
createFig_2_2_7910("#fig_2_2_10_ZWE",fig_2_2_10_data,"Zimbabwe",250,3e3,2);
});  

```
