---
title: "Section 4 TB prevention - Kendo UI charts"
author: "Hazim Timimi"
date: "`r Sys.Date()`"
knit: (function(inputFile, encoding) {
      out_dir <- "./html_drafts";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
output: 
  html_fragment:
    # Don’t include a table of contents
    toc: no
    # Set standard figure width to 12 inches
    fig_width: 12
    # Don’t write figure captions
    fig_caption: FALSE
    # Don't embed external resources (stylesheets, JS libraries) in the output 
    self_contained: FALSE

# To run this file and store output as html:
# rmarkdown::render(here::here("report/ch4_kendo.rmd"), output_file = "ch4_kendo.html", output_dir = here::here("report/html_drafts/"))
---

```{r setup, include=FALSE}
# Set chunk options.
# Results "asis" is useful to output markdown from a function
# Suppress messages, warnings and also the ## at the beginning of printed text

knitr::opts_chunk$set(echo = FALSE, 
                      results = "asis",
                      message = FALSE,
                      warning = FALSE,
                      error = TRUE)  # TEMP error=TRUE for debugging!

# Kill any attempt at using factors, unless we explicitly want them!
options(stringsAsFactors=FALSE)


# Load output packages ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

library(dplyr)
library(here)
library(jsonlite)

# Get the data sets and computed values/statistics for section 3.4 ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
source(here('report/ch4_prepare_data.r'))


```

<!-- Add links to stylesheets, JQuery and Kendo UI  -->

<link rel="stylesheet" href="https://kendo.cdn.telerik.com/2022.1.301/styles/kendo.common.min.css" />
<link rel="stylesheet" href="https://kendo.cdn.telerik.com/2022.1.301/styles/kendo.default.min.css" />
<link rel="stylesheet" href="https://kendo.cdn.telerik.com/2022.1.301/styles/kendo.default.mobile.min.css" />
<link rel="stylesheet" href="https://maxcdn.bootstrapcdn.com/bootstrap/3.4.1/css/bootstrap.min.css">

<script src="https://kendo.cdn.telerik.com/2022.1.301/js/jquery.min.js"></script>
<script src="https://kendo.cdn.telerik.com/2022.1.301/js/kendo.all.min.js"></script>
<script src="https://docs.telerik.com/kendo-ui/assets/bootstrap.min.js"></script>

<!-- End of links section -->


# 4. TB prevention

## Kendo UI charts


_Draft! Prepared `r Sys.Date()` using country-reported data CSV files dated `r csv_datestamp` and HIV TPT file dated `r csv_tpt_fix_datestamp`_

### Fig. 4.1</span> The global number of people provided with TB preventive treatment, 2015–`r report_year - 1`


<div id="fig_4_1"></div>


### <span class="red">Fig. 4.6</span> Completion of TB preventive treatment among contacts starting treatment,^a^ `r report_year - 2`


<div class="row">
<div class="col-md-4">
<div id="fig_4_6_afr"></div>
</div>
<div class="col-md-4">
<div id="fig_4_6_amr"></div>
</div>
<div class="col-md-4">
<div id="fig_4_6_sear"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_4_6_eur"></div>
</div>
<div class="col-md-4">
<div id="fig_4_6_emr"></div>
</div>
<div class="col-md-4">
<div id="fig_4_6_wpr"></div>
</div>
</div>



### <span class="red">Fig. 4.7</span> Provision of TB preventive treatment to people living with HIV^a^, 2005–`r report_year - 1`

<div id="fig_4_7_global"></div>

<div class="row">
<div class="col-md-4">
<div id="fig_4_7_afr"></div>
</div>
<div class="col-md-4">
<div id="fig_4_7_amr"></div>
</div>
<div class="col-md-4">
<div id="fig_4_7_sear"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_4_7_eur"></div>
</div>
<div class="col-md-4">
<div id="fig_4_7_emr"></div>
</div>
<div class="col-md-4">
<div id="fig_4_7_wpr"></div>
</div>
</div>


### <span class="red">Fig. 4.9</span> Completion of TB preventive treatment among people living with HIV^a^, `r report_year - 2`

<div id="fig_4_9"></div>


### <span class="red">Fig. 4.12</span> BCG vaccination coverage in infants^a^, globally and by WHO region 2019-2021


<div id="fig_4_12_global"></div>

<div class="row">
<div class="col-md-4">
<div id="fig_4_12_afr"></div>
</div>
<div class="col-md-4">
<div id="fig_4_12_amr"></div>
</div>
<div class="col-md-4">
<div id="fig_4_12_sear"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_4_12_eur"></div>
</div>
<div class="col-md-4">
<div id="fig_4_12_emr"></div>
</div>
<div class="col-md-4">
<div id="fig_4_12_wpr"></div>
</div>
</div>








<script type="text/javascript">

/* JSON data objects for the figures */

var fig_4_1_data = `r f4.1_data %>% pivot_wider(names_from = "TPT_category", id_cols = "year", values_from = "how_many") %>% toJSON("rows")`;


var fig_4_6_data = `r f4.6_data %>% mutate(size=10) %>%  select(-newinc_con_prevtx_cmplt) %>% toJSON("rows")`;


var fig_4_7_data = `r f4.7_data %>% toJSON("rows")`;


```{r, fig4.9_bubbles}


f4.9_bubbles <- f4.9_data %>% 
  # Convert regional names to colours
  mutate(col = case_when(
    entity == "African Region" ~ "red",
    entity == "Region of the Americas" ~ "orange",
    entity == "South-East Asia Region" ~ "yellow",
    entity == "European Region" ~ "green",
    entity == "Eastern Mediterranean Region" ~ "blue",
    entity == "Western Pacific Region" ~ "purple"),
    
    # Add a standard size for the bubbles
    size = 10) %>% 
  
  # Drop unnecessary variables
  select(-hiv_all_tpt_completed, -entity)

```


var fig_4_9_data = `r f4.9_bubbles %>%  toJSON("rows")`;


var fig_4_12_data = `r f4.12_data %>% toJSON("rows")`;

</script>



```{js, echo=FALSE}

/* Functions to create the figures */

function num_spacer(number) {
  // Change thousands separator to a non-breaking space as per WHO standard
  return kendo.toString(number, 'n0').replace(/,/g, '&nbsp;')
}

function axis_spacer(number) {
  // Change thousands separator to a non-breaking space as per WHO standard
      if (number >= 10 | number == 0 ) {
        return kendo.toString(number, 'n0').replace(/,/g, ' ');
      } else {return kendo.toString(number, 'n1').replace(/,/g, ' ');
             }
}

function createFig_4_1() {
        $("#fig_4_1").kendoChart({
            dataSource: {
                data: fig_4_1_data
            },
            legend: {
                position: "bottom"
            },
            seriesDefaults: {
                type: "column",
                stack: true,
                gap: 0.2,
                tooltip: {
                    visible: true,
                    template: "#= series.name #: (#= category #): #= num_spacer(value)#"
                }
            },
            series: [{
                name: "People living with HIV",
                field: "hiv_tpt",
                color: "#ffc425"
                }, {
                name: "Household contacts aged <5 years",
                field: "house_con04_tpt",
                color: "#9fcd25"
                }, {
                name: "Household contacts aged \u22655 years",
                field: "house_con5plus_tpt",
                color: "dark green"
            }],
            valueAxis: {
                labels: {
                    template: "#= kendo.format('{0}',value/1000000) #"
                },
                title: {
                    text: "Number of people (millions)"
                },
                line: {
                    visible: false
                }
            },
            categoryAxis: {
                field: "year",
                labels: {
                    rotation: "auto"
                },
                majorGridLines: {
                    visible: false
                }
            }
        });
}


function createFig_4_6(fig_ID, data, filter) {

  // Filter the dataset on the entity variable
  dataJSON = data.filter( element => element.entity == filter);

  $(fig_ID).kendoChart({
        dataSource: dataJSON,
        chartArea: {
            height: 250
        },  
        title: {
                  text: filter,
                  color: "black",
                  font: "bold 14px  Arial,Helvetica,sans-serif"
              },  
        legend: {
            // hide the legend
            visible: false
        },
        series: [{
        // Using bubble instead of scatter chart otherwise wouldn't be able to see 
        // the country name in the tooltip. So had to fix a bubble size in the data
        type: "bubble",
        xField: "newinc_con_prevtx",
        yField: "pct_completed",
        categoryField: "country",
        sizeField: "size",
        maxSize: 10,
        minSize: 10,
            }],
        yAxis: {
            labels: {
                template: "#= kendo.format('{0}', value) #%"
            },
            title: {
                text: "% contacts who completed\nTB preventive treatment",
                visible: true,
                font: "12px  Arial,Helvetica,sans-serif"
            },
            line: {
                visible: false
            }
        },
        xAxis: {
            type: "log",
            majorGridLines: {
                visible: false
            },
            labels: {
                template: "#= kendo.format('{0}',num_spacer(value)) #"
            },
            title: {
                text: "Contacts starting TB preventive treatment\n(log scale)",
                visible: true,
                font: "12px  Arial,Helvetica,sans-serif"
            }
        },
        tooltip: {
				    visible: true,
				    template: "#= category #, #= num_spacer(value.x)# started treatment, #= kendo.toString(value.y, 'n0')#% completed"
			}
    });
}



function createFig_4_7(fig_ID, data, filter) {

  // Filter the dataset on the entity variable
  dataJSON = data.filter( element => element.entity == filter);

  $(fig_ID).kendoChart({
        dataSource: dataJSON,
        chartArea: {
            height: 250
        },  
        title: {
                  text: filter,
                  color: "black",
                  font: "bold 14px  Arial,Helvetica,sans-serif"
              },  
        legend: {
            // hide the legend
            visible: false
        },
        seriesDefaults: {
            type: "line",
            tooltip: {
                visible: true,
                template: "#= series.name #: (#= category #): #= num_spacer(value)#"
            }
        },
        series: [{
            name: "Newly enrolled on HIV treatment",
            field: "hiv_tpt_new",
            color: "red",
    				dashType: "dash",
            markers: {
              size: 3
            }
            },{
            name: "Currently on HIV treatment",
            field: "hiv_tpt_all",
            color: "blue",  
            markers: {
              size: 3
            }
            }],
        valueAxis: {
            labels: {
                template: "#= axis_spacer(value/1e3) #"
            },
            title: {
                text: "Number of people (thousands)",
                visible: true,
                font: "12px  Arial,Helvetica,sans-serif"
            },
            line: {
                visible: false
            }
        },
        categoryAxis: {
            field: "year",
            labels: {
                step: 5
            },
            majorGridLines: {
                visible: false
            },
            title: {
                text: "Year",
                visible: true,
                font: "12px  Arial,Helvetica,sans-serif"
            }
        }
    });
}


function createFig_4_9() {

  $("#fig_4_9").kendoChart({
        dataSource:  {
            data: fig_4_9_data
        },
        legend: {
            position: "bottom"
        },
        chartArea: {
            height: 300
        }, 
        seriesDefaults: {
            // Using bubble instead of scatter chart otherwise wouldn't be able to see 
            // the country name in the tooltip. So had to fix a bubble size in the data
            type: "bubble",
            xField: "hiv_all_tpt_started",
            yField: "pct_completed",
            categoryField: "country",
            sizeField: "size",
            maxSize: 10,
            minSize: 10,
            colorField: "col"
        },
        // The individual series are here purely to create the legend
        series: [{
            name: "African Region",
            field: "entity",
            color: "red"
            }, {
            name: "Region of the Americas",
            field: "entity",
            color: "orange"
            }, {
            name: "South-East Asia Region",
            field: "entity",
            color: "yellow"
            }, {
            name: "European Region",
            field: "entity",
            color: "green"
            }, {
            name: "Eastern Mediterranean Region",
            field: "entity",
            color: "blue"
            }, {
            name: "Western Pacific Region",
            field: "entity",
            color: "purple"
            }],
        yAxis: {
            labels: {
                template: "#= kendo.format('{0}', value) #%"
            },
            title: {
                text: "% who completed TB preventive treatment",
                visible: true,
                font: "12px  Arial,Helvetica,sans-serif"
            },
            line: {
                visible: false
            }
        },
        xAxis: {
            type: "log",
            majorGridLines: {
                visible: false
            },
            labels: {
                template: "#= kendo.format('{0}',num_spacer(value)) #"
            },
            title: {
                text: "People living with HIV starting TB preventive treatment (log scale)",
                visible: true,
                font: "12px  Arial,Helvetica,sans-serif"
            }
        },
        tooltip: {
				    visible: true,
				    template: "#= category #, #= num_spacer(value.x)# started treatment, #= kendo.toString(value.y, 'n0')#% completed"
			}
    });
}

function createFig_4_12(fig_ID, data, filter) {
  
  		// Filter the dataset on the country variable
		dataJSON = data.filter( element => element.entity == filter);
  
		$(fig_ID).kendoChart({
			dataSource: dataJSON,
			chartArea: {
				height: 250
			},	
      title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},	
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "line"
			},
			series: [{
				field: "bcg_coverage",
				color: "#473c8b", 
        markers: {
          size: 3
        },
        tooltip: {
				visible: true,
				template: "#= category #: #= value.toPrecision(2) #%",
			}
			}],
			valueAxis: {
				labels: {
					format: "{0}"
				},
				title: {
					text: "BCG coverage (%)",
          font: "14px Arial,Helvetica,sans-serif"
				},
				line: {
					visible: false
				},
        max: 100,
        min: 0
			},
			categoryAxis: {
				field: "year",
        labels: {
					rotation: "auto"
          },
				majorGridLines: {
					visible: false
				},
				title: {
					text: "Year"
				}
			}

		});
}


```


```{js, echo=FALSE}

/* Create the figures after the document has been loaded */

$(document).ready(function() {
  
  createFig_4_1();
  
  createFig_4_6("#fig_4_6_afr", fig_4_6_data, "African Region");
  createFig_4_6("#fig_4_6_amr", fig_4_6_data, "Region of the Americas");
  createFig_4_6("#fig_4_6_sear", fig_4_6_data, "South-East Asia Region");
  createFig_4_6("#fig_4_6_eur", fig_4_6_data, "European Region");
  createFig_4_6("#fig_4_6_emr", fig_4_6_data, "Eastern Mediterranean Region");
  createFig_4_6("#fig_4_6_wpr", fig_4_6_data, "Western Pacific Region");
  
  
  createFig_4_7("#fig_4_7_global", fig_4_7_data, "Global");
  createFig_4_7("#fig_4_7_afr", fig_4_7_data, "African Region");
  createFig_4_7("#fig_4_7_amr", fig_4_7_data, "Region of the Americas");
  createFig_4_7("#fig_4_7_sear", fig_4_7_data, "South-East Asia Region");
  createFig_4_7("#fig_4_7_eur", fig_4_7_data, "European Region");
  createFig_4_7("#fig_4_7_emr", fig_4_7_data, "Eastern Mediterranean Region");
  createFig_4_7("#fig_4_7_wpr", fig_4_7_data, "Western Pacific Region");
  
  createFig_4_9();
  
  createFig_4_12("#fig_4_12_global", fig_4_12_data, "Global");
  createFig_4_12("#fig_4_12_afr", fig_4_12_data, "African Region");
  createFig_4_12("#fig_4_12_amr", fig_4_12_data, "Region of the Americas");
  createFig_4_12("#fig_4_12_sear", fig_4_12_data, "South-East Asia Region");
  createFig_4_12("#fig_4_12_eur", fig_4_12_data, "European Region");
  createFig_4_12("#fig_4_12_emr", fig_4_12_data, "Eastern Mediterranean Region");
  createFig_4_12("#fig_4_12_wpr", fig_4_12_data, "Western Pacific Region");

  
});

```


