---
title: "Chapter 6.2"
author: "Takuya Yamanaka" # Based on the test version of Chapter 5 "chapter_5_test" by Hazim 
date: "`r Sys.Date()`"
knit: (function(inputFile, encoding) {
      out_dir <- "./html_drafts";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
output: 
  html_fragment:
    # Don’t include a table of contents
    toc: no
    # Set standard figure width to 12 inches
    fig_width: 12
    # Don’t write figure captions
    fig_caption: FALSE
    css: columns.css
    # to save intermediate .png
    keep_md: true    
    # Don't embed external resources (stylesheets, JS libraries) in the output 
    self_contained: FALSE
   
---

```{r setup, include=FALSE}

# Set chunk options.
# Results "asis" is needed to avoid xtable output being html encoded
# Suppress messages, warnings and also the ## at the beginning of printed text
# Set a default figure width too.

knitr::opts_chunk$set(echo = FALSE,
                      results = "asis",
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 12)


# Load output packages ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
library(gtbreport)

library(jsonlite)
library(dplyr)
library(here)

# Get the data sets and computed values/statistics for section 6.1 ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
source(here('report/ch6-2_prepare_data.r'))

```

```{r setup2, include=FALSE}
# Chapter 1
# Set options, load data, utility functions
knitr::opts_chunk$set(echo = FALSE,
                      results = "asis",
                      message = FALSE,
                      warning = FALSE)

```

```{css, echo=FALSE}
/* Styles to make it easier to see in the html_fragment; this CSS can be included in the CSS widget of Sitefinity */
  

.red, .red a {
  color: #F21905; /* red text to show figure number */
}


```

<!-- Add links to stylesheets, JQuery and Kendo UI  -->

<link rel="stylesheet" href="https://kendo.cdn.telerik.com/2022.1.301/styles/kendo.common.min.css" />
<link rel="stylesheet" href="https://kendo.cdn.telerik.com/2022.1.301/styles/kendo.default.min.css" />
<link rel="stylesheet" href="https://kendo.cdn.telerik.com/2022.1.301/styles/kendo.default.mobile.min.css" />
<link rel="stylesheet" href="https://maxcdn.bootstrapcdn.com/bootstrap/3.4.1/css/bootstrap.min.css">

<script src="https://kendo.cdn.telerik.com/2022.1.301/js/jquery.min.js"></script>
<script src="https://kendo.cdn.telerik.com/2022.1.301/js/kendo.all.min.js"></script>
<script src="https://docs.telerik.com/kendo-ui/assets/bootstrap.min.js"></script>
<!-- End of links section -->


# 6.2 National surveys of costs faced by TB patients and their households   

_Draft! Prepared `r Sys.Date()` using country-reported data CSV files dated `r csv_datestamp`_  


### <span style="color:#F21905">Fig. 6.2.1</span> National surveys of costs faced by TB patients and their households since 2015: progress and plans (as of September `r report_year`)

<span class="red">**!static figure!**</span>



### <span style="color:#F21905">Fig. 6.2.2</span> Timeline of national TB patient cost surveys implemented since 2015 (as of September `r report_year`)


### <span style="color:#F21905">Fig. 6.2.3</span> Estimates of the percentage of TB patients and their households facing catastrophic costs,^a^ national surveys completed 2016&#8211;`r report_year` 

<div id="fig_6_2_3a"></div>

<div class="row">
<div class="col-md-6">
<div id="fig_6_2_3b"></div>
</div>
<div class="col-md-6">
<div id="fig_6_2_3c"></div>
</div>
</div>


### <span style="color:#F21905">Fig. 6.2.4</span> Distribution of costs faced by TB patients and their households in national surveys^a^ completed 2016&#8211;`r report_year`  

<div id="fig_6_2_4"></div>




### <span style="color:#F21905">Table 6.2.1</span> Status of social protection for people with TB, 30 high TB burden countries 

<span class="red">**!static figure!**</span>



<script type="text/javascript">
/* JSON data objects for the figures */

var fig_6_2_3a_data = `r f6.2.3a_data %>% filter(iso3!="AVE") %>% arrange(desc(catast_pct)) %>% add_row() %>% rbind.data.frame(filter(f6.2.3a_data, iso3=="AVE"))  %>% toJSON("rows")`   ;

var fig_6_2_3b_data = `r f6.2.3b_data %>% arrange(desc(factor(country, levels = f6.2.3_sel_order$country, ordered = TRUE)))  %>% filter(iso3!="AVE") %>% add_row() %>% rbind.data.frame(filter(f6.2.3b_data, iso3=="AVE"))  %>% toJSON("rows")`   ;

var fig_6_2_3c_data = `r f6.2.3c_data %>% arrange(desc(factor(country, levels = f6.2.3_sel_order$country, ordered = TRUE)))  %>% filter(iso3!="AVE") %>% add_row() %>% rbind.data.frame(filter(f6.2.3c_data, iso3=="AVE"))  %>% toJSON("rows")`   ;

var fig_6_2_4_data = `r f6.2.4_data %>% pivot_wider(names_from = cat, values_from = value) %>% arrange(desc(p_med)) %>% toJSON("rows")`   ;

</script>

```{js, echo=FALSE}

/* Functions to create the figures */


/* Functions to organize number with thousands separator */
function num_spacer(number) {
  // Change thousands separator to a non-breaking space as per WHO standard
      if (number >= 10 | number==0) {
        return kendo.toString(number, 'n0').replace(/,/g, '&nbsp;');
      } else if (number<10 & number>=1) {
      return kendo.toString(number, 'n1').replace(/,/g, '&nbsp;');
      } else {
      return kendo.toString(number, 'n2').replace(/,/g, '&nbsp;');
      }
}

function axis_spacer(number) {
  // Change thousands separator to a non-breaking space as per WHO standard
      if (number >= 10 | number == 0 ) {
        return kendo.toString(number, 'n0').replace(/,/g, ' ');
      } else {return kendo.toString(number, 'n1').replace(/,/g, ' ');
             }
}

/* format a number as thousands using GTB rounding rules */
function tb_format_thou(n) { 
  nt = n/1000; 
  if (nt < 100) {
    ntform = Number(nt.toPrecision(2))*1000;
      } else {
    ntform = Number(nt.toPrecision(3))*1000;
  } 
  return num_spacer(ntform)
}

function tb_format_pct(n) { 
  // Show percentages to 2 sig figs if under 100%
  if (n < 99.9) {
    return n.toPrecision(2)
      } else {
    return n.toPrecision(3)
  } 
}
  
/* Functions to create the figures */

function createFig_6_2_3(fig_ID, data, color, title) {
   
		$(fig_ID).kendoChart({
			dataSource: data,
			chartArea: {
				height: 900
			},	
			title: {
				text: title,
				color: "black",
				font: "bold 16px  Arial,Helvetica,sans-serif",
        align: "left"
			},	
			legend: {
				position: "bottom"
			},
			series: [{
        type: "bar",
				field: 0,
				opacity: 0
			}, {
        type: "line",
				field: "catast_pct",
        errorLowField: "catast_pct_lo",
        errorHighField: "catast_pct_hi",
        errorBars: {color: color, line: { width: 3 }},
        opacity: 0,
				color: color,
        markers: {
          visible: true,
          background: color,
          size: 10
        },
        
			tooltip: {
				visible: true,
        background: color,
				template: "#= category #: #= tb_format_pct(value)#%"
			}
			},{
				type: "rangeArea",
				fromField: "catast_pct_lo",
				toField: "catast_pct_hi",
				opacity: 0,
        color: color,
				tooltip: {
					visible: true,
          background: color,
				format: "{0}",
				template: "95% confidence interval (#= category #): #= tb_format_pct(value.from) #–#= tb_format_pct(value.to) #%"
				}
			}
              ],
			valueAxis: {
				labels: {
					format: "{0}"
				},
				title: {
					text: "Percentage (%)"
				},
				line: {
					visible: false
				},
        min: 0,
        max: 100,
        majorUnit: 25,
			},
			categoryAxis: {
				field: "country",
				labels: {
					rotation: "auto"
				},
				majorGridLines: {
					visible: true
				}
			}

		});
}


function createFig_6_2_4(fig_ID, data, y_axis_title_text) {
		$(fig_ID).kendoChart({
			dataSource: data,
			chartArea: {
				height: 600
			},	
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "bar",
        stack: {
          type: "100%"
        },
        gap: 0.2
			},
			series: [{
        name: "Direct medical",
				field: "p_med",
				color: "darkblue",
        tooltip: {
				visible: true,
				template: "Direct medical (#= category #): #= value.toPrecision(2) #%"
			}
			},{
        name: "Direct nonmedical",
				field: "p_nmed",
				color: "dodgerblue",
        tooltip: {
				visible: true,
				template: "Direct nonmedical (#= category #): #= value.toPrecision(2) #%"
			}
			},{
        name: "Indirect",
				field: "p_indirect",
				color: "goldenrod",
        tooltip: {
				visible: true,
				template: "Indirect (#= category #): #= value.toPrecision(2) #%"
			}
			},
              ],
			valueAxis: {

				title: {
					text: "Cost breakdown (%)"
				},
				line: {
					visible: false
				},
			},
			categoryAxis: {
				field: "country",
				title: {
					text: y_axis_title_text
				},
				labels: {
					rotation: "auto"
				},
				majorGridLines: {
					visible: false
				}			}
		});
}

```

```{js, echo=FALSE}

/* Create the figures after the document has been loaded */

$(document).ready(function () {
  createFig_6_2_3("#fig_6_2_3a",fig_6_2_3a_data,"dodgerblue","All TB");
  createFig_6_2_3("#fig_6_2_3b",fig_6_2_3b_data,"green","TB (first-line treatment)ᵇ");
  createFig_6_2_3("#fig_6_2_3c",fig_6_2_3c_data,"firebrick","Drug-resistant TBᵇ");
  createFig_6_2_4("#fig_6_2_4",fig_6_2_4_data,"");

 });  


```
