--- 
title: "Featured topic - MAF" 
author: "Monica Dias by Irwin Law" 
date: "`r Sys.Date()`" 
output:  
  html_fragment: 
    # Don’t include a table of contents 
    toc: no 
    # Set standard figure width to 12 inches 
    fig_width: 12 
    # Don’t write figure captions 
    fig_caption: FALSE 
    
    
# To run this file and store output as html:
# rmarkdown::render(here::here("report/ft_maf.rmd"), output_file = "ft_maf.html", output_dir = here::here("report/html_drafts/"))
--- 


```{r setup, include=FALSE} 
# Set options, load data, utility functions 

knitr::opts_chunk$set(echo = FALSE,  
                      results = "asis", 
                      message = FALSE, 
                      warning = FALSE) 
library(gtbreport)
library(here)
library(ggplot2)
library(readr)
library(scales)


# Quick functions for callouts in the text to the figures/tables:
# 1. lnk() Creates a link from text to a named anchor)
lnk <- function(display_name){
  return(paste0('<span class="red">[',
                display_name,
                '](#',
                gsub("[^a-zA-Z0-9]", "-", tolower(display_name)),
                ')</span>'))
}

# 2. anch() creates the named anchor to go above the figure or table
anch <- function(display_name){
  return(paste0('<a name="',
                gsub("[^a-zA-Z0-9]", "-", tolower(display_name)),
                '"></a>'))
}


``` 


```{css, echo=FALSE}

/* Styles to make it easier to see in the html_fragment; this CSS can be included in the CSS widget of Sitefinity */
  
.section.level3 {
  padding-bottom: 1em;
  border-bottom: 1px solid #BCBCBC;
  margin-bottom: 2em;
}


.subhead, .footnote {
  font-size: 80%;
  font-weight: normal;
  margin-top: 0;
  padding-top: 0;
}

.red, .red a {
  color: #F21905; /* red text to show figure number */
}


```



_Draft! Prepared `r Sys.Date()`_


### <span class="red">Fig. 1 </span>Trends on the public availability of annual national TB reports from 2020 to 2022
```{r fig_1, echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png',  fig.alt="Trends on the public availability of annual national TB reports from 2020 to 2022"}

#```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=7 }

maf_fig1<- read_csv('html_drafts/maf_fig1.csv')

p1<-ggplot(maf_fig1, aes(x = cat2, y = value, fill = year)) + 
  geom_bar(position="dodge", stat = "identity", width = 0.60)  +
  ylab("Percentage")  + 
  theme(axis.title.x = element_blank()) +
  scale_fill_discrete(
                         breaks=c("a", "b", "c"),
                         labels=c("2020", "2021", "2022"))+
  scale_x_discrete(labels = c('High TB burden countries',
                              'African Region',
                              'Region of the Americas',
                              'South-East Asia Region',
                              'European Region',
                              'Eastern Mediterranean Region',
                              'Western Pacific Region',
                              'Total'))+
  theme_gtb() +
  theme(axis.text = element_text(size = 8))  
p1

ggsave(here('report/html_drafts/fig1_maf.png'), width=12, height=5)


```


 <div class="footnote"> TB: tuberculosis; UN: United Nations; WHO: World Health Organization.
Source: UN General Assembly (2020) (_7_).</div> 



### <span class="red">Fig. 2 </span>Trends on the set up or strengthening of national multisectoral and multistakeholder accountability and review mechanisms, under high-level leadership available from 2020 to 2022
 
```{r fig_2.4.2, echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png',  fig.alt="Trends on the set up or strengthening of national multisectoral and multistakeholder accountability and review mechanisms, under high-level leadership available from 2020 to 2022"}

#```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=7 }

maf_fig2<- read_csv('html_drafts/maf_fig2.csv')

p2<-ggplot(maf_fig2, aes(x = cat2, y = value, fill = year)) + 
  geom_bar(position="dodge", stat = "identity", width = 0.60)  +
  ylab("Percentage")  + 
  theme(axis.title.x = element_blank()) +
  scale_fill_discrete(
                         breaks=c("a", "b", "c"),
                         labels=c("2020", "2021", "2022"))+
  scale_x_discrete(labels = c('High TB burden countries',
                              'African Region',
                              'Region of the Americas',
                              'South-East Asia Region',
                              'European Region',
                              'Eastern Mediterranean Region',
                              'Western Pacific Region',
                              'Total'))+
  theme_gtb() +
  theme(axis.text = element_text(size = 8))  
p2

ggsave(here('report/html_drafts/fig2_maf.png'), width=12, height=5)




```
 <div class="footnote"> CHW: community health worker; CS: civil society; CSO: civil society organization; MOH: ministry of health; TB: tuberculosis.</div> 


### <span class="red">Fig. 3 </span>Trends reported on the engagement of civil society and affected communities in the multisectoral accountability and review mechanism from 2020 to 2022
```{r fig_3, echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png',  fig.alt="Trends reported on the engagement of civil society and affected communities in the multisectoral accountability and review mechanism from 2020 to 2022"}

#```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=7 }

maf_fig3<- read_csv('html_drafts/maf_fig3.csv')

p3<-ggplot(maf_fig3, aes(x = cat2, y = value, fill = year)) + 
  geom_bar(position="dodge", stat = "identity", width = 0.60)  +
  ylab("Percentage")  + 
  theme(axis.title.x = element_blank()) +
  scale_fill_discrete(
                         breaks=c("a", "b", "c"),
                         labels=c("2020", "2021", "2022"))+
  scale_x_discrete(labels = c('High TB burden countries',
                              'African Region',
                              'Region of the Americas',
                              'South-East Asia Region',
                              'European Region',
                              'Eastern Mediterranean Region',
                              'Western Pacific Region',
                              'Total'))+
  theme_gtb() +
  theme(axis.text = element_text(size = 8))  +
  ylim(0, 100)
p3

ggsave(here('report/html_drafts/fig3_maf.png'), width=12, height=5)



```


